\subsection{Shader Module}
A Shader object represents one programmable stage of the graphics pipeline. A shader object has a type indicator, which tells whether the shader is a vertex or a fragment shader. The type of the shader is inferred when a GLSL shader file is loaded based on the extension of the shader file, when a '.vert' extension will be translated into a vertex shader and '.frag' into a fragment shader.

The shader module is also capable of compiling the GLSL source code into SPIR-V, an intermediate binary file format. The compilation is done using a third-party library, libshaderc, developed by Google. The SPIR-V code is then used to create the Vulkan shader module, an object used in the Vulkan pipeline creation later.

Shaders can define uniform objects containing various uniform variables. The uniforms names are checked for specific names and, if their types are as expected, their values are loaded automatically. This applies, for example, to a uniform named "modelViewProjection" of type "mat4". A similar process is adopted for vertex attributes, where their names and types are checked in order to load the correct values in the vertex buffer later on.

The uniform variables' names updated automatically are: "model", "view", "projection", "modelView", "modelViewProjection", all of their inverse matrices (with "Inverse" suffix), "cameraMatrix" (which is the view and projection transformations combined) and "normalMatrix" (the inverse transposed 3x3 modelView submatrix). All the matrices must be of type "mat4", except the normal matrix, which must be a "mat3" matrix.

The vertex attributes provided are: "position", "normal", "tangent" (each may be of types "vec3" or "vec4") and "uv" (texture coordinates, always "vec2"). Position vectors with four components will have W component equal to one; Normal vectors will have a W component equal to zero; Tangent vectors will have either 1 or -1 for its fourth component, indicating which side the binormal vector points to.

Shaders can also define image samplers of type "sampler2D" in GLSL. Image samplers will be initialized in the material module with a default white texture until the user selects a new one through the control window interface.

Uniforms objects and samplers must have different binding decoration numbers specified in GLSL using the layout qualifier.
